---
title: "Toronto Green Space"
output: html_notebook
---



```{r}
library(rsi)
library(terra)
library(tidyverse)
library(sf)
library(tmap)
library(here)
```

### Get locations and create neighborhood polygons
```{r}
Locations.df <- data.frame(Label=c("Church & Wellesley", "Southern Rosedale"), Latitude=c(43.665711, 43.676596), Longitude=c(-79.380943, -79.378481))
```

Convert to sf spatial object, define and change crs projection
```{r}
Locations.latlon <- sf::st_as_sf(Locations.df, coords=c("Longitude", "Latitude"),
                                crs=4326)
Locations.sf <- sf::st_transform(Locations.latlon, crs=32617)
```

Create a 500 m buffer around each location, then separate the buffers for the upcoming analysis

```{r}
Buffers <- sf::st_buffer(Locations.sf, dist=500)
CW <- Buffers[1,]
SR <- Buffers[2,]
```

Plot to check it
```{r}
tmap_mode("view")

tm_shape(Buffers) + tm_borders(col="red")
```

### Get Landsat8 images

Go online to find the dates you want for your imagery. \
\
Apply function get_landsat_imagery of R package rsi to each buffer (i.e., each local landscape) \
\
The resolution of Landsat8 is waaay too low (~30 m) for an urban setting, but it's free to download without an account or token, which is why we chose it for this tutorial.

```{r}
CW_imagery <- rsi::get_landsat_imagery(CW,
    start_date = "2024-08-14",
    end_date = "2024-08-15",
    output_filename = tempfile(fileext = ".tif"))

SR_imagery <- rsi::get_landsat_imagery(SR,
    start_date = "2024-08-14",
    end_date = "2024-08-15",
    output_filename = tempfile(fileext = ".tif"))
```

Convert to raster
```{r}
CW_imagery <- terra::rast(CW_imagery)
SR_imagery <- terra::rast(SR_imagery)
```

Mask to area of interest
```{r}
CW_imagery <- terra::mask(CW_imagery, CW)
SR_imagery <- terra::mask(SR_imagery, SR)
```

Plot to check
```{r}
par(mfrow=c(1,2))
terra::plotRGB(CW_imagery, r = 4, g = 3, b = 2, stretch = "lin")
terra::plotRGB(SR_imagery, r = 4, g = 3, b = 2, stretch = "lin")
```

Write the files out in a rasterstack so you don't have to re-download it later
```{r}
writeRaster(CW_imagery, paste0(here::here(),"/data/CW_landsat8_14Aug2024.tif"),
   overwrite=TRUE)
writeRaster(SR_imagery, paste0(here::here(),"/data/SR_landsat8_14Aug2024.tif"),
   overwrite=TRUE)
```

Plot in infrared to see the greenspaces better
```{r}
par(mfrow=c(1,2))
terra::plotRGB(CW_imagery, r = "N", g = "R", b = "G", stretch = "lin")
terra::plotRGB(SR_imagery, r = "N", g = "R", b = "G", stretch = "lin")

```
### Calculate NDVI
```{r}
CW_ndvi <- (CW_imagery[["N"]] - CW_imagery[["R"]])/
                    (CW_imagery[["N"]] + CW_imagery[["R"]])

SR_ndvi <- (SR_imagery[["N"]] - SR_imagery[["R"]])/
                    (SR_imagery[["N"]] + SR_imagery[["R"]])
```


```{r}
par(mfrow=c(1,2))
terra::plot(CW_ndvi, range = c(-1, 1))
terra::plot(SR_ndvi, range = c(-1, 1))
```
Calculate the mean NDVI for each neighborhood
```{r}
data.frame(
           CW_NDVI=mean(terra::values(CW_ndvi), na.rm=TRUE),
           SR_NDVI=mean(terra::values(SR_ndvi), na.rm=TRUE))
```
```{r}
threshold = 0.6
CW_veg <- as.numeric(CW_ndvi > threshold)
SR_veg <- as.numeric(SR_ndvi > threshold)
```


```{r}
par(mfrow=c(1,2))
terra::plot(CW_veg, range = c(-1, 1), col=c("grey", "darkgreen"))
terra::plot(SR_veg, range = c(-1, 1), col=c("grey", "darkgreen"))
```

```{r}
data.frame(
           PctVeg_CW=mean(terra::values(CW_veg), na.rm=TRUE),
           PctVeg_SR=mean(terra::values(SR_veg), na.rm=TRUE))
```



